---
title: "Australia Numbats"
author: "Helia Tehrani"
date: "2023-03-11"
output: 
  prettydoc::html_pretty:
    theme: tactile
    toc: TRUE
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../Output/")
```                     
 
 ## Introduction
 Welcome to Tidy Tuesday for this week. I will be working with numbat data that was gathered in Australia this week. Because they are so adorable, despite being an endangered species, people are enamored with them. I'm trying to respond to the following queries:
 .Where in Australia can you locate numbats?
 .– Which season is more common, summer or winter?
 ![Numbat Pic](https://th.bing.com/th/id/OIP.vJlALNXB1o8urBaTxWkojAHaE7?pid=ImgDet&rs=1)
 
 ## Libraries
```{r Libraries}
library(ggmap)
library(tidyverse)
library(here)
library(ggsn)
library(emojifont)
library(readr)
library(nationalparkcolors)
library(praise)
library(devtools)
library(beyonce)
library(ggthemes)
library(hrbrthemes)
library(ggridges)
library(nationalparkcolors)
library(tidytuesdayR)
library(ggplot2)
library(viridis)
library(lubridate)
```
 # Data
The data can be manually entered using the code.
```{r}
numbats <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-03-07/numbats.csv')
```
#To display a map of Australia and the locations of the numerous Numbat sightings, I utilized the code below from Google Maps. The map demonstrates that there have been numerous sightings in southwest Australia.
```{r Numbat Map}
Australia <- get_map("Australia", zoom = 6, maptype = "terrain") #satellite map of australia
ggmap(Australia) + #creat map
  geom_point(data = numbats, #uses data for points
             aes(x = decimalLongitude, #uses longitude from numbat data
                 y = decimalLatitude)) #uses latitude from numbat data
```
#I plotted sightings on the maps using the raw data. Let's now purge the data so we can respond to the second question.
```{r Sightings in Different Seasons}
clean_numbats <- numbats %>% #numbats data
  drop_na(month) %>% #drop na's
  select(decimalLatitude, decimalLongitude, month, eventDate) %>% #The columns I want to use
  mutate(Month = month(eventDate), #extract month out of UTC column
         season  = case_when( # make a season
           Month %in% c(12,1,2) ~ "winter",
           Month %in% c(3,4,5) ~ "spring",
           Month %in% c(6,7,8) ~ "summer",
           Month %in% c(9,10,11) ~ "fall"),
         season = factor(season), #make season a factor
         season = fct_relevel(season, c("winter","spring","summer","fall"))) #arrange as seasons fall through the year
ggplot(clean_numbats,
       aes(x = Month,
           y = season,
           fill = season))+
  geom_density_ridges(alpha = 0.8)+ #ridges plot
  theme_ridges()+
  theme(legend.position = "none")+ #deletes legend 
  labs(y = "Season",
       title = "Number of Sightings of Numbats per Season")
       
```       