---
title: "Soccer Game Info"
author: "Helia Tehrani"
date: "2023-04-12"
output: 
  prettydoc::html_pretty:
    theme: tactile
    toc: TRUE
---
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../Output/")
```                  
 
#Intro:
*soccer dataset that examines various teams' victories and losses is part of this week's tidy tuesday. I want to use some new software I've learned to try to produce some amusing graphs using this random dataset. In order to become more accustomed to these fundamentals of coding, I also want to practice tidying data with mutate, select, and if else.*
    
## Load libraries 
```{r}
library(tidyverse)
library(here)
library(magick)
library(PNWColors)
library(patchwork)
library(GGally)
library(viridis)
library(pals)
```
## Read Data 
```{r }
soccer <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-04-04/soccer21-22.csv')
```


```{r}
halftime <- table(soccer$HTR, exclude = NULL) ## Ir's in R base
prop.table(table(soccer$HTR, exclude = NULL)) ## demonstrates the percentage of datasets where each category is present. 
fulltime <- table(soccer$FTR, exclude=NULL)
fulltime
soccer_home <- soccer %>%
  group_by(HomeTeam) %>% 
  summarize(sum_homegoals=sum(FTHG), 
            sum_shots=sum(HS)) 
soccer_away <- soccer %>% 
  group_by(AwayTeam) %>% 
  summarize(sum_awaygoals=sum(FTAG), 
            sum_shots=sum(AS))
```

## Plot the Data 
The first figure will show the number of shots taken as a function of total shots from full-time scores.

```{r, goals_and_shots}
ggplot(data=soccer_home, 
       aes(x=sum_homegoals, 
           y=sum_shots)) + 
  geom_point() + 
  labs(x="Total Goals from the Winning Games", 
       y="Total Shots Taken", 
       title="Goals Scored vs Goals on Goals for Home Teams") + 
  theme_classic() +
  theme(plot.title=element_text(hjust = 0.5)) 
  
ggsave(here("soccer Game Info", "Outputs", "goals_shots.jpg"), 
       width=10, height=8)
```

## Plotting the fouls 
In order to determine whether a team is more aggressive when playing at home or away, this plot will show the red cards awarded to each team both at home and away.

```{r, redcardsplot}
p1 <- ggplot(data=soccer, 
       aes(y=HomeTeam, 
           x=HR, 
           fill=HomeTeam)) + 
  geom_col() + 
  theme_bw() + 
  labs(x="The Red Cards of the Home Team", 
       y="Home Team", 
       title="Number of Red Cards per Home Team", 
       fill="Home Team") + 
  theme(plot.title=element_text(hjust = 0.6)) 
p1
p2 <- ggplot(data=soccer,
       aes(x=AR, 
           y=AwayTeam, 
           fill=AwayTeam)) + 
  geom_col() + 
  theme_bw() + 
  labs(x="Away Team Red Cards", 
       y="Away Team", 
       title="Number of Red Cards per Away Team", 
       fill="Away Team") + 
  theme(plot.title=element_text(hjust = 0.5)) 
p2
p1 + p2  
  ggsave(here("Soccer Game Info", "Outputs", "redcards.jpg"), 
         width=9, height=6)
```
#Conclusion
*On tidy Tuesday, I looked into the soccer data collection, which contained a wealth of data on soccer games, who won, shots on goal, halftime and final scores, and fouls.
Because that "H" and "A" were characters rather than numeric variables, I was unable to determine how to tally up the total number of victories on each occasion when they competed as home team vs. away team.
Even so, I'm glad I had the opportunity to practice organizing dataframes, adding new columns, and learning more about patchwork.